---
title: "AGP Age Prediction"
output: 
  html_document:    
    toc: true
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(reticulate)
# Change if needed
use_condaenv("trac")
library(tidyverse)
library(phyloseq)
library(trac)
library(knitr)
library(kableExtra)
```

# Age analysis with trac and two-stage procedure

The aim of this document is to detect potential log-ratio pairs of possibly aggregated microbiome (compositional) data with additional non-compositional covariates that are predictive of age based on the American Gut Project data obtained from the reproducible repo of the trac paper.

## Preprocess data


```{r}
data <- readRDS("D:/KUSS/School/Biostat seminar/AGP_BMI_cat/data/AGP_10.RDS")
data_subset <- data.frame(data@sam_data)  
```


Extract a host related variables with subjects not having diabetes or have taken antibiotics the last year.


```{r}
levels_str <- c("Never", 
                "Rarely (a few times/month)",
                "Occasionally (1-2 times/week)",
                "Regularly (3-5 times/week)",
                "Daily")
level_str_food <- c("Never", 
                "Rarely (less than once/week)",
                "Occasionally (1-2 times/week)",
                "Regularly (3-5 times/week)",
                "Daily")
levels_sleep <- c("Less than 5 hours", 
                 "5-6 hours",
                 "6-7 hours",
                 "7-8 hours",
                 "8 or more hours")
levels_bmi <- c("Underweight",
                "Normal",
                "Overweight",
                "Obese")

relevant_features_1 <- data_subset %>%
  select(bmi_cat,
         alcohol_consumption,
         exercise_frequency,
         sleep_duration,
         fruit_frequency,
         vegetable_frequency,
         age_years,
         sex,
         antibiotic_history,
         diabetes,
         non_food_allergies_unspecified,
         allergic_to_i_have_no_food_allergies_that_i_know_of,
         dog,
         flossing_frequency,
         frozen_dessert_frequency,
         last_travel,
         last_move,
         milk_substitute_frequency,
         nail_biter,
         meat_eggs_frequency,
         olive_oil,
         one_liter_of_water_a_day_frequency,
         prepared_meals_frequency,
         probiotic_frequency,
         red_meat_frequency,
         seafood_frequency,
         smoking_frequency,
         sugar_sweetened_drink_frequency,
         sugary_sweets_frequency,
         teethbrushing_frequency,
         vitamin_b_supplement_frequency,
         vitamin_d_supplement_frequency,
         whole_eggs,
         whole_grain_frequency) %>%
  mutate(across(.cols = everything(), as.character)) %>%
  filter(antibiotic_history == "I have not taken antibiotics in the past year.") %>%
  filter(diabetes == "I do not have this condition") %>%
  select(-antibiotic_history, -diabetes) %>%
  mutate(across(where(is.character), ~ str_replace_all(., "Unspecified", replacement = ""))) %>%
  mutate(across(where(is.character), ~ str_replace_all(., "unknown", replacement = ""))) %>%
  mutate(across(where(is.character), ~ str_replace_all(., "yes", replacement = "true"))) %>%
  mutate(across(where(is.character), ~ str_replace_all(., "no", replacement = "false"))) %>%
  mutate(across(where(is.character), ~ str_replace_all(., "Yes", replacement = "true"))) %>%
  mutate(across(where(is.character), ~ str_replace_all(., "No", replacement = "false"))) %>%
  mutate(across(where(is.character), ~ na_if(.,""))) 
na_count <- sapply(relevant_features_1, function(x) sum(length(which(is.na(x)))))
na_count <- names(na_count[na_count > 300])
relevant_features_1 <- relevant_features_1 %>%
select(-all_of(na_count)) %>%
  na.omit() %>%
  mutate(age_years = as.numeric(age_years))

check_distinct_values <- relevant_features_1 %>% 
  mutate(across(everything(), n_distinct))
binary_names <- colnames(relevant_features_1)[which(check_distinct_values[1, ] == 2)]
binary_names <- binary_names[binary_names != "sex"]

food_levels <- c("frozen_dessert_frequency", "one_liter_of_water_a_day_frequency", 
           "meat_eggs_frequency", "milk_substitute_frequency", "olive_oil",
           "prepared_meals_frequency", "red_meat_frequency", "seafood_frequency",
           "sugar_sweetened_drink_frequency",
           "sugary_sweets_frequency", "whole_eggs", "whole_grain_frequency",
           "fruit_frequency", "vegetable_frequency")
normal_levels <- c("probiotic_frequency", "teethbrushing_frequency", 
           "vitamin_b_supplement_frequency", "vitamin_d_supplement_frequency", 
           "smoking_frequency", "exercise_frequency", "flossing_frequency"
           )

food_levels <- food_levels[!food_levels %in% na_count]
normal_levels <- normal_levels[!normal_levels %in% na_count]

relevant_features_2 <- relevant_features_1 %>%
  mutate(bmi_cat = as.numeric(factor(bmi_cat, levels = levels_bmi)),
         bmi_cat = ifelse(is.na(bmi_cat), 0, bmi_cat),
         sleep_duration = as.numeric(factor(sleep_duration,
                                     levels = levels_sleep)),
         sex = as.numeric(sex == "female"),
         last_travel = as.numeric(last_travel == "I have falset been outside of my country of residence in the past year.")
        ) %>%
  mutate(across(.cols = all_of(binary_names), ~ as.numeric(.x == "true"))) %>%
  mutate(across(.cols = all_of(normal_levels), ~  as.numeric(factor(.x, levels = levels_str))))

relevant_features <- relevant_features_2
```

Create subset of the original phyloseq object to analyze and root in bacteria


```{r}
subset_logical <- rownames(relevant_features)
data_subset <- prune_samples(subset_logical, data)
agp <- phyloseq(otu_table(data_subset),
                        tax_table(data_subset),
                        sample_data(relevant_features))
agp <- subset_taxa(agp , Rank1 == "k__Bacteria")
agp
```


## Data processing for predicitive performance

Preprocessing for trac, creating the A matrix

```{r}
y <- (sample_data(agp)$age_years)
tax <- agp@tax_table@.Data
# add an OTU column
tax <- cbind(tax, rownames(tax))
colnames(tax)[ncol(tax)] <- "OTU"

# make it so labels are unique
for (i in seq(2, 7)) {
  # add a number when the type is unknown... e.g. "g__"
  ii <- nchar(tax[, i]) == 3
  if (sum(ii) > 0)
    tax[ii, i] <- paste0(tax[ii, i], 1:sum(ii))
}
# cumulative labels are harder to read but easier to work with:
for (i in 2:8) {
  tax[, i] <- paste(tax[, i-1], tax[, i], sep = "::")
}
tax <- as.data.frame(tax, stringsAsFactors = TRUE)

# form phylo object:
tree1 <- tax_table_to_phylo(~Rank1/Rank2/Rank3/Rank4/Rank5/Rank6/Rank7/OTU,
                            data = tax, collapse = TRUE)

# convert this to an A matrix to be used for aggregation:
A <- phylo_to_A(tree1)

dat <- list(y = y,
            x = t(agp@otu_table@.Data),
            tree = tree1,
            tax = tax,
            A = A,
            sample_data = as_tibble(sample_data(agp)))
```


## Model fitting

Create train test split and pseudo-count of 1

### trac

```{r}
set.seed(123)
additional_covariates <- data.frame(dat$sample_data) %>%
  select(-age_years)
```

```{r}
set.seed(123)
log_pseudo <- function(x, pseudo_count = 1) log(x + pseudo_count)
z <- log_pseudo(dat$x)

y <- dat$y
summary(y)

ntot <- length(y)
n <- round(0.8 * ntot)
tr <- sample(ntot, n)

# define training and test data
ytr <- y[tr]
yte <- y[-tr]

ztr <- z[tr, ]
zte <- z[-tr, ]
additional_covariates_tr <- additional_covariates[tr, ]
additional_covariates_te <- additional_covariates[-tr, ]
w_add_cov <- rep(1, ncol(additional_covariates))
```

```{r}

table(ytr)
table(yte)

```

Fit model

```{r}
reticulate::py_run_string("import numpy as np; np.infty = np.inf")

fit <- trac(ztr,
            ytr, 
            A = dat$A, 
            additional_covariates = additional_covariates_tr, 
            method = "regr", 
            w_additional_covariates = w_add_cov,
            normalized = FALSE)

plot_trac_path <- function(fit, iw = 1, coef = c("alpha", "beta", "gamma")) {
  coef <- match.arg(coef)
  fit[[iw]][[coef]] %>% 
    as.matrix() %>% 
    t() %>%
    tibble::as_tibble() %>% 
    dplyr::mutate(frac = fit[[iw]]$fraclist) %>%
    tidyr::pivot_longer(cols = -frac) %>% 
    ggplot2::ggplot(ggplot2::aes(x = frac, y = value, 
                                 group = name)) + 
    ggplot2::geom_line(alpha = 0.2) + 
    ggplot2::scale_x_log10() + 
    ggplot2::theme(legend.position = "none") + 
    ggplot2::labs(y = coef, x = "Fraction of lambda_max") +
    theme_bw() +
    ylab("Alpha")
}

fit_plot <- plot_trac_path(fit)
fit_plot
```


```{r}
w_add_cov_05 <- rep(0.5, ncol(additional_covariates))
```


### trac with less penalization

```{r}
fit_wo_pen <- trac(ztr, 
            ytr, 
            A = dat$A,
            additional_covariates = additional_covariates_tr, 
            method = "regr", 
            w_additional_covariates = w_add_cov_05,
            normalized = FALSE)


fit_plot_wo_pen <- plot_trac_path(fit_wo_pen)
fit_plot_wo_pen
```

### slc


```{r}
# fit slc 

fit_slc <- sparse_log_contrast(ztr, 
            ytr, 
            additional_covariates = additional_covariates_tr, 
            method = "regr", 
            w_additional_covariates = w_add_cov,
            normalized = FALSE)

plot_slc_path <- function(fit, iw = 1, coef = c("beta")) {
  coef <- match.arg(coef)
  fit[[coef]] %>% 
    as.matrix() %>% 
    t() %>%
    tibble::as_tibble() %>% 
    dplyr::mutate(frac = fit$fraclist) %>%
    tidyr::pivot_longer(cols = -.data$frac) %>% 
    ggplot2::ggplot(ggplot2::aes(x = frac, y = value, 
                                 group = name)) + 
    ggplot2::geom_line(alpha = 0.2) + 
    ggplot2::scale_x_log10() + 
    ggplot2::theme(legend.position = "none") + 
    ggplot2::labs(y = coef, x = "Fraction of lambda_max") +
    theme_bw() +
    ylab("beta")
}

fit_plot_slc <- plot_slc_path(fit_slc)
fit_plot_slc
```



### slc with less penalization

```{r}
# fit slc 

fit_slc_wo_pen <- sparse_log_contrast(ztr, 
            ytr, 
            additional_covariates = additional_covariates_tr, 
            method = "regr", 
            w_additional_covariates = w_add_cov_05,
            normalized = FALSE)


fit_plot_slc_wo_pen <- plot_slc_path(fit_slc_wo_pen)
fit_plot_slc_wo_pen
```


### CV for hyperparameter tuning


```{r}
set.seed(123)
folds <- ggb:::make_folds(n = length(ytr), nfolds = 5)
cvfit <- cv_trac(fit, Z = ztr, y = ytr, A = dat$A, 
                 additional_covariates = additional_covariates_tr,
                 summary_function = base::mean,
                 folds = folds)
```


```{r}
cv_trac_plot <- data.frame(non_zero = cvfit$cv[[1]]$nonzeros,
           error = cvfit$cv[[1]]$m,
           se = cvfit$cv[[1]]$se) %>%
  ggplot(aes(x = non_zero, y = error)) + 
  geom_line(size = 1) +
  geom_errorbar(aes(ymin = error - se, ymax = error + se), width = 20, size = 0.8,
                color = "grey50") +
  geom_point(size = 2.5, alpha = 1) +
  geom_vline(xintercept = cvfit$cv[[1]]$nonzeros[cvfit$cv[[1]]$i1se], linetype = 2,
             size = 0.8) +
  geom_vline(xintercept = cvfit$cv[[1]]$nonzeros[cvfit$cv[[1]]$ibest], 
             size = 0.8) +
  theme_bw() +
  xlab("Number of nonzero alpha") +
  ylab("Cross-validation Error")

  
cv_trac_plot
```


```{r}
set.seed(123)
cvfit_wo_pen <- cv_trac(fit_wo_pen, Z = ztr, y = ytr, A = dat$A, 
                 additional_covariates = additional_covariates_tr,
                 summary_function = base::mean,
                 folds = folds)

```


```{r}
cv_trac_plot_wo_pen <- data.frame(non_zero = cvfit_wo_pen$cv[[1]]$nonzeros,
           error = cvfit_wo_pen$cv[[1]]$m,
           se = cvfit_wo_pen$cv[[1]]$se) %>%
  ggplot(aes(x = non_zero, y = error)) + 
  geom_line(size = 1) +
  geom_errorbar(aes(ymin = error - se, ymax = error + se), width = 20, size = 0.8,
                color = "grey50") +
  geom_point(size = 2.5, alpha = 1) +
  geom_vline(xintercept = cvfit_wo_pen$cv[[1]]$nonzeros[cvfit_wo_pen$cv[[1]]$i1se], linetype = 2,
             size = 0.8) +
  geom_vline(xintercept = cvfit_wo_pen$cv[[1]]$nonzeros[cvfit_wo_pen$cv[[1]]$ibest], 
             size = 0.8) +
  theme_bw() +
  xlab("Number of nonzero alpha") +
  ylab("Cross-validation Error") 

  
cv_trac_plot_wo_pen
```

```{r}
set.seed(123)
cvfit_slc <- cv_sparse_log_contrast(fit_slc, Z = ztr, y = ytr,
                 additional_covariates = additional_covariates_tr,
                 summary_function = base::mean,
                 folds = folds)
```



```{r}
cv_slc_plot <- data.frame(non_zero = cvfit_slc$cv$nonzeros,
           error = cvfit_slc$cv$m,
           se = cvfit_slc$cv$se) %>%
  ggplot(aes(x = non_zero, y = error)) + 
  geom_line(size = 1) +
  geom_errorbar(aes(ymin = error - se, ymax = error + se), width = 20, size = 0.8,
                color = "grey50") +
  geom_point(size = 2.5, alpha = 1) +
  geom_vline(xintercept = cvfit_slc$cv$nonzeros[cvfit_slc$cv$i1se], linetype = 2,
             size = 0.8) +
  geom_vline(xintercept = cvfit_slc$cv$nonzeros[cvfit_slc$cv$ibest], 
             size = 0.8) +
  theme_bw() +
  xlab("Number of nonzero alpha") +
  ylab("Cross-validation Error") 
  
cv_slc_plot
```


```{r}
set.seed(123)
cvfit_slc_wo_pen <- cv_sparse_log_contrast(fit_slc_wo_pen, Z = ztr, y = ytr,
                 additional_covariates = additional_covariates_tr,
                 summary_function = base::mean,
                 folds = folds)
```



```{r}
cv_slc_plot_wo_pen <- data.frame(non_zero = cvfit_slc_wo_pen$cv$nonzeros,
           error = cvfit_slc_wo_pen$cv$m,
           se = cvfit_slc_wo_pen$cv$se) %>%
  ggplot(aes(x = non_zero, y = error)) + 
  geom_line(size = 1) +
  geom_errorbar(aes(ymin = error - se, ymax = error + se), width = 20, size = 0.8,
                color = "grey50") +
  geom_point(size = 2.5, alpha = 1) +
  geom_vline(xintercept = cvfit_slc_wo_pen$cv$nonzeros[cvfit_slc_wo_pen$cv$i1se], linetype = 2,
             size = 0.8) +
  geom_vline(xintercept = cvfit_slc_wo_pen$cv$nonzeros[cvfit_slc_wo_pen$cv$ibest], 
             size = 0.8) +
  theme_bw() +
  xlab("Number of nonzero alpha") +
  ylab("Cross-validation Error")
  
cv_slc_plot_wo_pen
```

### Prediction and non-zero


```{r}
y_hat_trac <- predict_trac(fit = fit, new_Z = zte,
             new_additional_covariates = additional_covariates_te)
y_hat_trac_wo_pen <- predict_trac(fit = fit_wo_pen, new_Z = zte,
             new_additional_covariates = additional_covariates_te)
y_hat_slc <- predict_sparse_log_contrast(fit = fit_slc,
                                  new_Z = zte,
                                  new_additional_covariates =
                                    additional_covariates_te)
y_hat_slc_wo_pen <- predict_sparse_log_contrast(fit = fit_slc_wo_pen, 
                                                new_Z = zte,
                                                new_additional_covariates =
                                                  additional_covariates_te)


rmse <- function(actual, predicted) {
  sqrt(mean((actual - predicted)^2))
}

# 1SE

y_hat_trac_1SE <- y_hat_trac[[1]][, cvfit$cv[[1]]$i1se]
y_hat_trac_wo_pen_1SE <- y_hat_trac_wo_pen[[1]][, cvfit_wo_pen$cv[[1]]$i1se]
y_hat_slc_1SE <- y_hat_slc[, cvfit_slc$cv$i1se]
y_hat_slc_wo_pen_1SE <- y_hat_slc_wo_pen[, cvfit_slc_wo_pen$cv$i1se]



rmse_trac_1SE <- rmse(yte, y_hat_trac_1SE)
rmse_trac_wo_pen_1SE <- rmse(yte, y_hat_trac_wo_pen_1SE)
rmse_slc_1SE <- rmse(yte, y_hat_slc_1SE)
rmse_slc_wo_pen_1SE <- rmse(yte, y_hat_slc_wo_pen_1SE)

nz_1SE <- (fit[[1]]$alpha[, cvfit$cv[[1]]$i1se])
nz_1SE <- nz_1SE[abs(nz_1SE) > 0]
nz_n_1SE <- length(nz_1SE)

nz_wo_pen_1SE <- (fit_wo_pen[[1]]$alpha[, cvfit_wo_pen$cv[[1]]$i1se])
nz_wo_pen_1SE <- nz_wo_pen_1SE[abs(nz_wo_pen_1SE) > 0]
nz_wo_pen_n_1SE <- length(nz_wo_pen_1SE)

nz_slc_1SE <- (fit_slc$beta[, cvfit_slc$cv$i1se])
nz_slc_1SE <- nz_slc_1SE[abs(nz_slc_1SE) > 0]
nz_slc_n_1SE <- length(nz_slc_1SE)

nz_slc_wo_pen_1SE <- (fit_slc_wo_pen$beta[, cvfit_slc_wo_pen$cv$i1se])
nz_slc_wo_pen_1SE <- nz_slc_wo_pen_1SE[abs(nz_slc_wo_pen_1SE) > 0]
nz_slc_wo_pen_n_1SE <- length(nz_slc_wo_pen_1SE)


# Best

y_hat_trac <- y_hat_trac[[1]][, cvfit$cv[[1]]$ibest]
y_hat_trac_wo_pen <- y_hat_trac_wo_pen[[1]][, cvfit_wo_pen$cv[[1]]$ibest]
y_hat_slc <- y_hat_slc[, cvfit_slc$cv$ibest]
y_hat_slc_wo_pen <- y_hat_slc_wo_pen[, cvfit_slc_wo_pen$cv$ibest]

rmse_trac <- rmse(yte, y_hat_trac)
rmse_trac_wo_pen <- rmse(yte, y_hat_trac_wo_pen)
rmse_slc <- rmse(yte, y_hat_slc)
rmse_slc_wo_pen <- rmse(yte, y_hat_slc_wo_pen)

nz <- (fit[[1]]$alpha[, cvfit$cv[[1]]$ibest])
nz <- nz[abs(nz) > 0]
nz_n <- length(nz)

nz_wo_pen <- (fit_wo_pen[[1]]$alpha[, cvfit_wo_pen$cv[[1]]$ibest])
nz_wo_pen <- nz_wo_pen[abs(nz_wo_pen) > 0]
nz_wo_pen_n <- length(nz_wo_pen)

nz_slc <- (fit_slc$beta[, cvfit_slc$cv$ibest])
nz_slc <- nz_slc[abs(nz_slc) > 0]
nz_slc_n <- length(nz_slc)

nz_slc_wo_pen <- (fit_slc_wo_pen$beta[, cvfit_slc_wo_pen$cv$ibest])
nz_slc_wo_pen <- nz_slc_wo_pen[abs(nz_slc_wo_pen) > 0]
nz_slc_wo_pen_n <- length(nz_slc_wo_pen)

```

```{r}
rmse_trac_1SE 
rmse_trac_wo_pen_1SE
rmse_slc_1SE 
rmse_slc_wo_pen_1SE

rmse_trac 
rmse_trac_wo_pen
rmse_slc 
rmse_slc_wo_pen
```


```{r}

# get n for compositional and non compositional selected features
# trac based methods have :: in the names for compositional and
# slc based methods have numbers for compositional features

get_n_selected_compostional_non_compositional <- function(nz, trac = TRUE) {
  if (is.null(names(nz))) {
    c(comp = 0, non_comp = 0)
  } else {
    if(trac) {
      distinct_comp_non_comp <- names(nz) %>% str_count(pattern = "::")
      c(comp = sum(distinct_comp_non_comp >= 1), 
           non_comp = sum(distinct_comp_non_comp == 0))
    } else {
      distinct_comp_non_comp <- names(nz) %>% str_detect(pattern = "[[:digit:]]")
      c(comp = sum(distinct_comp_non_comp == TRUE), 
           non_comp = sum(distinct_comp_non_comp == FALSE))
    }
  }
}


get_n_selected_compostional_non_compositional(nz)
get_n_selected_compostional_non_compositional(nz_wo_pen)
get_n_selected_compostional_non_compositional(nz_slc, trac = FALSE)
get_n_selected_compostional_non_compositional(nz_slc_wo_pen, trac = FALSE)
```

```{r}

# get n for compositional and non compositional selected features
# trac based methods have :: in the names for compositional and
# slc based methods have numbers for compositional features

get_selected_non_compositional <- function(nz, trac = TRUE) {
  if (is.null(names(nz))) {
    NULL
  } else {
    if(trac) {
      distinct_comp_non_comp <- names(nz) %>% str_count(pattern = "::")
      names(nz)[distinct_comp_non_comp == 0]
    } else {
      distinct_comp_non_comp <- names(nz) %>% str_detect(pattern = "[[:digit:]]")
           names(nz)[distinct_comp_non_comp == FALSE]
    }
  }
}


# get distinct taxa number by 1. splitting the log-ratios and counting the
# distinct taxa
get_distinct_taxa <- function(nz, trac = TRUE) {
  nz %>%
    enframe() %>%
    separate(col = name, into = c("first", "second"), sep = "///") %>%
    filter(!is.na(second)) %>%
    pivot_longer(cols = c("first", "second"), values_to = "taxa") %>%
    distinct(taxa) %>%
    nrow()
}
```


### Second Stage

#### trac


```{r}
set.seed(123)
get_selected_non_compositional(nz)
two_stage <- second_stage(Z = ztr, A = dat$A, y = ytr, 
                     additional_covariates = 
                       additional_covariates_tr[
                         , get_selected_non_compositional(nz), drop = FALSE],
                     method = "regr", 
                     betas = fit[[1]]$alpha[, cvfit$cv[[1]]$ibest],
                     criterion = "1se", topk = 300, alpha = 0,
                     nfolds = length(folds)  
                     #, classo = FALSE,  w_additional = 1
                     )
nz_second <- two_stage$log_ratios[two_stage$log_ratios != 0]
nz_second_n <- length(nz_second)

y_hat <- predict_second_stage(fit = two_stage, new_Z = zte, 
                              new_additional_covariates = 
                                additional_covariates_te[ 
                                  ,get_selected_non_compositional(nz),
                                  drop = FALSE])
rmse_second <- rmse(yte, y_hat)
rmse_second
```


```{r}
set.seed(123)
get_selected_non_compositional(nz)
two_stage_wo_pen <- second_stage(Z = ztr, A = dat$A, y = ytr, 
                     additional_covariates = 
                       additional_covariates_tr[
                         , get_selected_non_compositional(nz), drop = FALSE],
                     method = "regr", 
                     betas = fit[[1]]$alpha[, cvfit$cv[[1]]$ibest],
                     criterion = "1se", topk = 300, alpha = 0,
                     #, w_additional = 0.5, 
                     nfolds = length(folds)
                     )

nz_second_wo_pen <- two_stage_wo_pen$log_ratios[two_stage_wo_pen$log_ratios != 0]
nz_second_wo_pen_n <- length(nz_second_wo_pen)

y_hat_wo_pen <- predict_second_stage(fit = two_stage_wo_pen, new_Z = zte, 
                              new_additional_covariates = 
                                additional_covariates_te[ 
                                  ,get_selected_non_compositional(nz), 
                                  drop = FALSE])
rmse_second_wo_pen <- rmse(yte, y_hat_wo_pen)
rmse_second_wo_pen
```


#### trac wo pen


```{r}

set.seed(123)
get_selected_non_compositional(nz_wo_pen)
two_stage_wo_pen_normal <-
  second_stage(Z = ztr, A = dat$A, y = ytr,
               additional_covariates =
                 additional_covariates_tr[ 
                   , get_selected_non_compositional(nz_wo_pen), drop = FALSE],
               method = "regr",
               betas = fit_wo_pen[[1]]$alpha[, cvfit_wo_pen$cv[[1]]$ibest],
               criterion = "1se", topk = 300, alpha = 0,# w_additional = 1, 
               nfolds = length(folds)
               )

nz_second_wo_pen_normal <- two_stage_wo_pen_normal$log_ratios[two_stage_wo_pen_normal$log_ratios != 0]
nz_second_wo_pen_normal_n <- length(nz_second_wo_pen_normal)
nz_second_wo_pen_normal_n
y_hat_wo_pen_normal <-
  predict_second_stage(fit = two_stage_wo_pen_normal, new_Z = zte,
                       new_additional_covariates =
                         additional_covariates_te[ 
                           , get_selected_non_compositional(nz_wo_pen), 
                           drop = FALSE])
rmse_second_wo_pen_normal <- rmse(yte, y_hat_wo_pen_normal)
rmse_second_wo_pen_normal

```


```{r}

set.seed(123)
two_stage_wo_pen_wo_pen <- second_stage(Z = ztr, A = dat$A, y = ytr,
                     additional_covariates =
                       additional_covariates_tr[ 
                         , get_selected_non_compositional(nz_wo_pen), 
                         drop = FALSE],
                     method = "regr",
                     betas = fit_wo_pen[[1]]$alpha[, cvfit_wo_pen$cv[[1]]$ibest],
                     criterion = "1se", alpha = 0, #w_additional = 0.5, 
                     nfolds = length(folds)
                     )
nz_second_wo_pen_wo_pen <- two_stage_wo_pen_wo_pen$log_ratios[two_stage_wo_pen_wo_pen$log_ratios != 0]
nz_second_wo_pen_wo_pen_n <- length(nz_second_wo_pen_wo_pen)

y_hat_wo_pen_wo_penn <-
  predict_second_stage(fit = two_stage_wo_pen_wo_pen, new_Z = zte,
                       new_additional_covariates =
                         additional_covariates_te[ ,
                                                   get_selected_non_compositional(nz_wo_pen), drop = FALSE])
rmse_second_wo_pen_wo_pen <- rmse(yte, y_hat_wo_pen_wo_penn)
rmse_second_wo_pen_wo_pen

```

#### slc

```{r}

set.seed(123)
two_stage_slc <- second_stage(Z = ztr, y = ytr, 
                     additional_covariates = 
                       additional_covariates_tr[
                         , get_selected_non_compositional(nz_slc, trac = FALSE),
                         drop = FALSE],
                     method = "regr", 
                     betas = fit_slc$beta[, cvfit_slc$cv$ibest],
                     criterion = "1se", alpha = 0,
                     nfolds = length(folds)#,  w_additional = 1
                     )


nz_second_slc <- two_stage_slc$log_ratios[two_stage_slc$log_ratios != 0]
nz_second_slc_n <- length(nz_second_slc)

y_hat_slc_two <- predict_second_stage(fit = two_stage_slc, new_Z = zte, 
                              new_additional_covariates = 
                                additional_covariates_te[ ,get_selected_non_compositional(nz_slc, trac = FALSE), 
                                                     drop = FALSE])
rmse_second_slc <- rmse(yte, y_hat_slc_two[, 1])
rmse_second_slc

```


```{r}

set.seed(123)
two_stage_slc_wo_pen <- second_stage(Z = ztr, y = ytr, 
                     additional_covariates = 
                       additional_covariates_tr[ , get_selected_non_compositional(nz_slc, trac = FALSE), drop = FALSE],
                     method = "regr", 
                     betas = fit_slc$beta[, cvfit_slc$cv$ibest],
                     criterion = "1se", alpha = 0, #w_additional = 0.5, 
                     nfolds = length(folds)
                     )

nz_second_slc_wo_pen <- two_stage_slc_wo_pen$log_ratios[two_stage_slc_wo_pen$log_ratios != 0]
nz_second_slc_wo_pen_n <- length(nz_second_slc_wo_pen)

y_hat_slc_two_wo_pen <- predict_second_stage(fit = two_stage_slc_wo_pen, new_Z = zte, 
                              new_additional_covariates = additional_covariates_te[ , get_selected_non_compositional(nz_slc, trac = FALSE), 
                                                                                    drop = FALSE])
rmse_second_slc_wo_pen <- rmse(yte, y_hat_slc_two_wo_pen[, 1])
rmse_second_slc_wo_pen

```


#### slc wo pen


```{r}

set.seed(123)
two_stage_slc_wo_pen_normal <-
  second_stage(Z = ztr, y = ytr,
               additional_covariates =
                 additional_covariates_tr[ , get_selected_non_compositional(nz_slc_wo_pen, trac = FALSE), drop = FALSE],
               method = "regr",
               betas = fit_slc_wo_pen$beta[, cvfit_slc_wo_pen$cv$ibest],
               criterion = "1se", alpha = 0,# w_additional = 1,
                     nfolds = length(folds)
                     )

nz_second_slc_wo_pen_normal <- 
  two_stage_slc_wo_pen_normal$log_ratios[two_stage_slc_wo_pen_normal$log_ratios != 0]
nz_second_slc_wo_pen_normal_n <- length(nz_second_slc_wo_pen_normal)

y_hat_slc_wo_pen_normal <-
  predict_second_stage(fit = two_stage_slc_wo_pen_normal, new_Z = zte,
                       new_additional_covariates =
                         additional_covariates_te[ , get_selected_non_compositional(nz_slc_wo_pen, trac = FALSE), drop = FALSE])
rmse_second_slc_wo_pen_normal <- rmse(yte, y_hat_slc_wo_pen_normal[, 1])
rmse_second_slc_wo_pen_normal
```


```{r}

set.seed(123)
two_stage_slc_wo_pen_wo_pen <- second_stage(Z = ztr, y = ytr,
                     additional_covariates =
                       additional_covariates_tr[ , get_selected_non_compositional(nz_slc_wo_pen, trac = FALSE), drop = FALSE],
                     method = "regr",
                     betas = fit_slc_wo_pen$beta[, cvfit_slc_wo_pen$cv$ibest],
                     criterion = "1se", topk = 300, alpha = 0, #w_additional = 0.5,
                     nfolds = length(folds)
                     )

nz_second_slc_wo_pen_wo_pen <- 
  two_stage_slc_wo_pen_wo_pen$log_ratios[two_stage_slc_wo_pen_wo_pen$log_ratios != 0]
nz_second_slc_wo_pen_wo_pen_n <- length(nz_second_slc_wo_pen_wo_pen)

y_hat_slc_wo_pen_wo_penn <-
  predict_second_stage(fit = two_stage_slc_wo_pen_wo_pen, new_Z = zte,
                       new_additional_covariates =
                         additional_covariates_te[ ,
                                                  get_selected_non_compositional(nz_slc_wo_pen, trac = FALSE), drop = FALSE])
rmse_second_slc_wo_pen_wo_pen <- rmse(yte, y_hat_slc_wo_pen_wo_penn[, 1])
rmse_second_slc_wo_pen_wo_pen
```


## Baseline Models: only Microbiome


### trac

```{r}
fit_baseline <- trac(ztr,
            ytr, 
            A = dat$A, 
            method = "regr")

fit_plot_baseline <- plot_trac_path(fit_baseline)
fit_plot_baseline
```

### slc

```{r}
fit_slc_baseline <- sparse_log_contrast(ztr, 
            ytr, 
            method = "regr")


fit_plot_slc_baseline <- plot_slc_path(fit_slc_baseline)
fit_plot_slc_baseline
```

### CV for hyperparameter

```{r}
set.seed(123)
cvfit_baseline <- cv_trac(fit_baseline, Z = ztr, y = ytr, A = dat$A, 
                 summary_function = base::mean,
                 folds = folds)


nz_baseline <- (fit_baseline[[1]]$alpha[, cvfit_baseline$cv[[1]]$ibest])
nz_baseline <- nz_baseline[abs(nz_baseline) > 0]
nz_baseline_n <- length(nz_baseline)

y_hat_trac_baseline <- predict_trac(fit = fit_baseline, new_Z = zte)

y_hat_trac_baseline <- y_hat_trac_baseline[[1]][, cvfit_baseline$cv[[1]]$ibest]

rmse_trac_baseline <- rmse(yte, y_hat_trac_baseline)
rmse_trac_baseline
```



```{r}
set.seed(123)
cvfit_slc_baseline <- cv_sparse_log_contrast(fit_slc_baseline, Z = ztr, y = ytr,
                 summary_function = base::mean,
                 folds = folds)


nz_slc_baseline <- (fit_slc_baseline$beta[, cvfit_slc_baseline$cv$ibest])
nz_slc_baseline <- nz_slc_baseline[abs(nz_slc_baseline) > 0]
nz_slc_baseline_n <- length(nz_slc_baseline)


y_hat_slc_baseline <- predict_sparse_log_contrast(fit = fit_slc_baseline,
                                  new_Z = zte)
y_hat_slc_baseline <- y_hat_slc_baseline[, cvfit_slc_baseline$cv$ibest]

rmse_slc_baseline <- rmse(yte, y_hat_slc_baseline)
rmse_slc_baseline
```

### Second Stage baseline

#### trac

```{r}
set.seed(123)
two_stage_baseline <- second_stage(Z = ztr, A = dat$A, y = ytr,
                     method = "regr",
                     betas = fit_baseline[[1]]$alpha[, cvfit_baseline$cv[[1]]$ibest],
                     criterion = "1se", alpha = 0,
                     nfolds = length(folds)
                     )

nz_second_baseline <- two_stage_baseline$log_ratios[two_stage_baseline$log_ratios != 0]
nz_second_baseline_n <- length(nz_second_wo_pen_normal)

y_hat_baseline <- predict_second_stage(fit = two_stage_baseline, new_Z = zte)
rmse_second_baseline <- rmse(yte, y_hat_baseline)
rmse_second_baseline
```

#### SLC

```{r}
set.seed(123)
two_stage_slc_baseline <- second_stage(Z = ztr, y = ytr, 
                     method = "regr", 
                     betas = fit_slc_baseline$beta[, cvfit_slc_baseline$cv$ibest],
                     criterion = "1se", alpha = 0,
                     nfolds = length(folds)
                     )

nz_second_slc_baseline <- two_stage_slc_baseline$log_ratios[two_stage_slc_baseline$log_ratios != 0]
nz_second_slc_baseline_n <- length(nz_second_slc_baseline)

y_hat_slc_two_baseline <- predict_second_stage(fit = two_stage_slc_baseline, new_Z = zte)

rmse_second_slc_baseline <- rmse(yte, y_hat_slc_two_baseline)
rmse_second_slc_baseline
```


## Baseline Models: only non-compositional


```{r}
non_comp <- cbind(ytr, additional_covariates_tr)
fit_lm <- glm(ytr ~ ., data = non_comp, family = "gaussian")
y_hat_lm <- predict(fit_lm, newdata = additional_covariates_te)

n_lm <- ncol(additional_covariates_tr)
rmse_lm <- rmse(yte, y_hat_lm)
rmse_lm
```

```{r}
set.seed(123)
fit_sparse <- glmnet::cv.glmnet(as.matrix(additional_covariates_tr), ytr,
                                family = "gaussian",
                                nfolds = 5,  type.measure = "mse",
                                alpha = 1)
y_hat_sparse <- predict(fit_sparse, newx = as.matrix(additional_covariates_te))
nz_sparse <- as.matrix(coef(fit_sparse))[-1, ]
nz_sparse <- nz_sparse[nz_sparse != 0]
n_sparse <- sum(as.matrix(coef(fit_sparse)) != 0) - 1
rmse_sparse <- rmse(yte, y_hat_sparse)
rmse_sparse
```

## Calculate Pearson

### Pearson

```{r}
cor_logistic <- cor(y_hat_lm, yte)
cor_lasso <- cor(y_hat_sparse, yte)

cor_slc_baseline <- cor(y_hat_slc_baseline, yte)
cor_trac_baseline <- cor(y_hat_trac_baseline, yte)
cor_second_baseline <- cor(y_hat_baseline, yte)
cor_second_slc_baseline <- cor(y_hat_slc_two_baseline, yte)

cor_trac <- cor(y_hat_trac, yte)
cor_trac_wo_pen <- cor(y_hat_trac_wo_pen, yte)
cor_slc <- cor(y_hat_slc, yte)
cor_slc_wo_pen <- cor(y_hat_slc_wo_pen, yte)

cor_second_normal_normal <- cor(y_hat, yte)
cor_second_normal_wo_pen <- cor(y_hat_wo_pen, yte)
cor_second_wo_pen_normal <- cor(y_hat_wo_pen_normal, yte)
cor_second_wo_pen_wo_pen <- cor(y_hat_wo_pen_wo_penn, yte)

cor_slc_second_normal_normal <- cor(y_hat_slc_two, yte)
cor_slc_second_normal_wo_pen <- cor(y_hat_slc_two_wo_pen, yte)
cor_slc_second_wo_pen_normal <- cor(y_hat_slc_wo_pen_normal, yte)
cor_slc_second_wo_pen_wo_pen <- cor(y_hat_slc_wo_pen_wo_penn, yte)
```


```{r}
# Get predictions and actual values
predictions <- y_hat_wo_pen_normal
actual <- yte

# Calculate regression metrics
results <- data.frame(
  RMSE = sqrt(mean((actual - predictions)^2)),
  MAE = mean(abs(actual - predictions)),
  R_squared = cor(actual, predictions)^2,
  Pearson_cor = cor(actual, predictions)
)
print(results)
```
```{r}
caret::postResample(pred = y_hat_wo_pen_normal, obs = yte)
```

## Save models

```{r}
models_list <- list(lm_baseline = fit_lm,
                    lasso_baseline = fit_sparse,
                    slc_baseline = list(fit_slc_baseline, cvfit_slc_baseline),
                    slc_baseline_second = two_stage_slc_baseline,
                    trac_baseline  = list(fit_baseline, cvfit_baseline),
                    trac_baseline_second = two_stage_baseline,
                    slc = list(fit_slc, cvfit_slc),
                    slc_wo_pen = list(fit_slc_wo_pen, cvfit_slc_wo_pen),
                    trac = list(fit, cvfit),
                    trac_wo_pen = list(fit_wo_pen, cvfit_wo_pen),
                    slc_second = two_stage_slc,
                    slc_second_wo_pen = two_stage_slc_wo_pen,
                    slc_wo_pen_second = two_stage_slc_wo_pen_normal,
                    slc_wo_pen_second_wo_pen = two_stage_slc_wo_pen_wo_pen,
                    trac_second = two_stage,
                    trac_second_wo_pen = two_stage_wo_pen,
                    trac_wo_pen_second = two_stage_wo_pen_normal,
                    trac_wo_pen_second_wo_pen = two_stage_wo_pen_wo_pen)
saveRDS(models_list, file = "../AGP_age/models/models.RDS")
```



## Create figures and tables

### tables


```{r}
df_summary <- data.frame(
  Method = c("Logistic \n Regression", "Sparse \n Logistic \n Regression", "SLC", "SLC + \n second stage",
             "trac", "trac + \n second stage", "SLC", "SLC",
             "trac", "trac", rep("SLC + \n second stage", 4),
             rep("trac + \n second stage", 4)),
  Microbiome = c(rep("-", 2), rep("+", 16)),
  Meta = c(rep("+", 2), rep("-", 4), rep("+", 12)),
  Stage_I = c(rep("-", 6), rep(c(1, 0.5), 2), rep(c(1, 1, 0.5, 0.5), 2)),
  Stage_II = c(rep("-", 10), rep(c(1, 0.5), 4)),
  c_compositional = c("-",
               "-",
               nz_slc_baseline_n,
               get_distinct_taxa(nz_second_slc_baseline),
               nz_baseline_n,
               get_distinct_taxa(nz_second_baseline),
               get_n_selected_compostional_non_compositional(nz_slc, trac = FALSE)[1],
               get_n_selected_compostional_non_compositional(nz_slc_wo_pen, trac = FALSE)[1],
               get_n_selected_compostional_non_compositional(nz)[1],
               get_n_selected_compostional_non_compositional(nz_wo_pen)[1],
               get_distinct_taxa(nz_second_slc)[1],
               get_distinct_taxa(nz_second_slc_wo_pen)[1],
               get_distinct_taxa(nz_second_slc_wo_pen_normal)[1],
               get_distinct_taxa(nz_second_slc_wo_pen_wo_pen)[1],
               get_distinct_taxa(nz_second)[1],
               get_distinct_taxa(nz_second_wo_pen)[1],
               get_distinct_taxa(nz_second_wo_pen_normal)[1],
               get_distinct_taxa(nz_second_wo_pen_wo_pen)[1]),
  c_non_compositional = c(n_lm,
                          n_sparse,
                          "-",
                          "-",
                          "-",
                          "-",
               get_n_selected_compostional_non_compositional(nz_slc, trac = FALSE)[2],
               get_n_selected_compostional_non_compositional(nz_slc_wo_pen, trac = FALSE)[2],
               get_n_selected_compostional_non_compositional(nz)[2],
               get_n_selected_compostional_non_compositional(nz_wo_pen)[2],
               get_n_selected_compostional_non_compositional(nz_second_slc, trac = FALSE)[2],
               get_n_selected_compostional_non_compositional(nz_second_slc_wo_pen, trac = FALSE)[2],
               get_n_selected_compostional_non_compositional(nz_second_slc_wo_pen_normal, trac = FALSE)[2],
               get_n_selected_compostional_non_compositional(nz_second_slc_wo_pen_wo_pen, trac = FALSE)[2],
               get_n_selected_compostional_non_compositional(nz_second)[2],
               get_n_selected_compostional_non_compositional(nz_second_wo_pen)[2],
               get_n_selected_compostional_non_compositional(nz_second_wo_pen_normal)[2],
               get_n_selected_compostional_non_compositional(nz_second_wo_pen_wo_pen)[2]),
  c_log_ratio = c(rep("-", 3),
                  nz_second_slc_baseline_n,
                  "-",
                  nz_second_baseline_n,
                  rep("-", 4),
                  get_n_selected_compostional_non_compositional(nz_second_slc, trac = FALSE)[1],
               get_n_selected_compostional_non_compositional(nz_second_slc_wo_pen, trac = FALSE)[1],
               get_n_selected_compostional_non_compositional(nz_second_slc_wo_pen_normal, trac = FALSE)[1],
               get_n_selected_compostional_non_compositional(nz_second_slc_wo_pen_wo_pen, trac = FALSE)[1],
               get_n_selected_compostional_non_compositional(nz_second)[1],
               get_n_selected_compostional_non_compositional(nz_second_wo_pen)[1],
               get_n_selected_compostional_non_compositional(nz_second_wo_pen_normal)[1],
               get_n_selected_compostional_non_compositional(nz_second_wo_pen_wo_pen)[1]),
  RMSE = c(rmse_lm,
          rmse_sparse,
          rmse_slc_baseline,
          rmse_second_slc_baseline,
          rmse_trac_baseline,
          rmse_second_baseline,
          rmse_slc,
          rmse_slc_wo_pen,
          rmse_trac,
          rmse_trac_wo_pen,
          rmse_second_slc,
          rmse_second_slc_wo_pen,
          rmse_second_slc_wo_pen_normal,
          rmse_second_slc_wo_pen_wo_pen,
          rmse_second,
          rmse_second_wo_pen,
          rmse_second_wo_pen_normal,
          rmse_second_wo_pen_wo_pen),
    Pearson = c(cor_logistic,
          cor_lasso,
          cor_slc_baseline,
          cor_second_slc_baseline,
          cor_trac_baseline,
          cor_second_baseline,
          cor_slc,
          cor_slc_wo_pen,
          cor_trac,
          cor_trac_wo_pen,
          cor_slc_second_normal_normal,
          cor_slc_second_normal_wo_pen,
          cor_slc_second_wo_pen_normal,
          cor_slc_second_wo_pen_wo_pen,
          cor_second_normal_normal,
          cor_second_normal_wo_pen,
          cor_second_wo_pen_normal,
          cor_second_wo_pen_wo_pen)
  ) %>%
  mutate(RMSE = round(RMSE, 3),
         Pearson = round(Pearson, 3))

df_summary 

df_summary %>%
  mutate(Method = linebreak(Method, align = "l")) %>%
  kbl("latex", digits = c(rep(1,4), rep(3,6)), 
      col.names = c("Model","Bacteria","Meta", "Stage I", "Stage II", "Taxa", "Meta", "Log-ratio", "AUC", "MCC"), booktabs = TRUE,
      align = c("l", rep("c", 8)), escape = FALSE) %>%
     add_header_above(c(" " = 1, "Data type" = 2, "Meta weight" = 2, "# selected" = 3)) %>%
  save_kable(
    "../AGP_age/results/summary_models.tex") 
```





### Figures

```{r}
trac_coeff_plot <- nz_second_wo_pen_wo_pen %>% enframe() %>%
  mutate(name = str_replace_all(name, "'", "")) %>%
  mutate(non_comp = case_when(str_starts(name, "k__") ~ 1,
                              TRUE ~ 0)) %>%
  group_by(non_comp) %>%
  slice_max(order_by = abs(value), n = 5) %>%
  ungroup() %>%
  mutate(name = str_replace(name, "k__Bacteria::", ""),
           name = str_replace(name, "Bacteria::", ""),
           name = str_remove_all(name, "[a-z]__"),
           name = str_replace(name, "sex", "female"),
           name = str_replace_all(name, "_", " "),
           name = str_replace_all(name, "///", "\n")) %>%
  arrange(value) %>%
  mutate(name = factor(name, name)) %>%
  ggplot(aes(x = value, y = name)) +
  geom_point(size = 2) +
  geom_segment(aes(x = 0, xend = value, y = name, yend = name)) +
  xlab("Coefficients") +
  ylab("") +
  theme_bw() +
  theme(axis.text.y = element_text(size = rel(0.8)))
ggsave(filename = "../AGP_age/results/agp_coeff_trac.png", plot = trac_coeff_plot)
```


```{r}
OTU_name <- enframe(as.character(tax$OTU)) %>%
  separate(value, sep = "::", into = c("Kingdom", "Phylum", "Class",
                                       "Order", "Family", "Genus",
                                       "Species", "OTU"),
           remove = FALSE) %>%
  mutate(name = OTU) %>%
  select(name, value) %>%
  rename(names = value)

OTU_slc_second <- nz_second_slc_wo_pen_normal %>% 
  enframe() %>%
  separate(col = name, into = c("first", "second"), sep = "///") %>%
  filter(!is.na(second)) %>%
  rename(name = first) %>%
  left_join(OTU_name, by = "name") %>%
  rename(first = name,
         name = second,
         names_first = names) %>%
  left_join(OTU_name, by = "name") %>%
  rename(second = name,
         names_second = names) %>%
  mutate(name = paste0(first, "///", second),
         names = paste0(names_first, "///", names_second)) %>%
  select(name, names)


slc_coeff_plot <- nz_second_slc_wo_pen_normal %>% enframe() %>%
  left_join(OTU_slc_second, by = "name") %>%
  mutate(names = case_when(is.na(names) ~ name,
                           TRUE ~ names)) %>%
    mutate(name = str_replace_all(names, "'", ""),
           name = str_replace(name, "k__Bacteria::", ""),
           name = str_remove_all(name, "[a-z]__"),
           name = str_replace(name, "sex", "female"),
           name = str_replace_all(name, "_", " "),
           name = str_replace_all(name, "///", "\n")) %>%
  arrange(value) %>%
  mutate(name = factor(name, name)) %>%
  ggplot(aes(x = value, y = name)) +
  geom_point(size = 2) +
  geom_segment(aes(x = 0, xend = value, y = name, yend = name)) +
  xlab("Coefficients") +
  ylab("") +
  theme_bw() +
  theme(axis.text.y = element_text(size = rel(0.8)))
ggsave(filename = "../AGP_age/results/agp_coeff_slc.png", plot = slc_coeff_plot)
```


```{r}
# check which of the non-compositional covaritates have been selected by all
# methods

# get logical vector if the non zero features are non-compositional for plotting
# the intersect between the different methods
get_non_compositional_trac <- function(nz) {
  distinct_comp_non_comp <- names(nz) %>% str_count(pattern = "::")
  distinct_comp_non_comp == 0
}

get_non_compositional_slc <- function(nz) {
  distinct_comp_non_comp <- names(nz) %>% str_detect(pattern = "[[:digit:]]")
  distinct_comp_non_comp != TRUE
}


non_comp_trac <- names(nz[get_non_compositional_trac(nz)])
names(non_comp_trac) <- rep("trac (w = 1)", length(non_comp_trac))

non_comp_trac_wo_pen <- names(nz_wo_pen[get_non_compositional_trac(nz_wo_pen)])
names(non_comp_trac_wo_pen) <- rep("trac (w = 0.5)", length(non_comp_trac_wo_pen))

non_comp_slc <- names(nz_slc[get_non_compositional_slc(nz_slc)])
names(non_comp_slc) <- rep("slc (w = 1)", length(non_comp_slc))

non_comp_slc_wp_pen <- names(nz_slc_wo_pen[get_non_compositional_slc(nz_slc_wo_pen)])
names(non_comp_slc_wp_pen) <- rep("slc (w = 0.5)", length(non_comp_slc_wp_pen))

lasso_selected <- names(nz_sparse)
names(lasso_selected) <- rep("lasso", length(lasso_selected))



non_comp_trac_second <- names(nz_second[get_non_compositional_trac(nz_second)])
names(non_comp_trac_second) <- rep("trac (w = 1) + second (w = 1)", length(non_comp_trac_second))

non_comp_trac_second_wo_pen <- names(nz_second_wo_pen[get_non_compositional_trac(nz_second_wo_pen)])
names(non_comp_trac_second_wo_pen) <- rep("trac (w = 1) + second (w = 0.5)", length(non_comp_trac_second_wo_pen))

non_comp_trac_second_wo_pen_normal <- names(nz_second_wo_pen_normal[get_non_compositional_trac(nz_second_wo_pen_normal)])
names(non_comp_trac_second_wo_pen_normal) <- rep("trac (w = 0.5) + second (w = 1)", length(non_comp_trac_second_wo_pen_normal))

non_comp_trac_second_wo_pen_wo_pen <- names(nz_second_wo_pen_wo_pen[get_non_compositional_trac(nz_second_wo_pen_wo_pen)])
names(non_comp_trac_second_wo_pen_wo_pen) <- rep("trac (w = 0.5) + second (w = 0.5)", length(non_comp_trac_second_wo_pen_wo_pen))


non_comp_slc_second <- names(nz_second_slc[get_non_compositional_slc(nz_second_slc)])
names(non_comp_slc_second) <- rep("slc (w = 1) + second (w = 1)", length(non_comp_slc_second))

non_comp_slc_second_wo_pen <- names(nz_second_slc_wo_pen[get_non_compositional_slc(nz_second_slc_wo_pen)])
names(non_comp_slc_second_wo_pen) <- rep("slc (w = 1) + second (w = 0.5)", length(non_comp_slc_second_wo_pen))

non_comp_slc_second_wo_pen_normal <- names(nz_second_slc_wo_pen_normal[get_non_compositional_slc(nz_second_slc_wo_pen_normal)])
names(non_comp_slc_second_wo_pen_normal) <- rep("slc (w = 0.5) + second (w = 1)", length(non_comp_slc_second_wo_pen_normal))

non_comp_slc_second_wo_pen_wo_pen <- names(nz_second_slc_wo_pen_wo_pen[get_non_compositional_slc(nz_second_slc_wo_pen_wo_pen)])
names(non_comp_slc_second_wo_pen_wo_pen) <- rep("slc (w = 0.5), second (w = 0.5)", length(non_comp_slc_second_wo_pen_wo_pen))

df_joint_selected <- bind_rows(enframe(non_comp_trac),
          enframe(non_comp_trac_wo_pen),
          enframe(non_comp_slc),
          enframe(non_comp_slc_wp_pen),
          enframe(lasso_selected),
          enframe(non_comp_trac_second),
           enframe(non_comp_trac_second_wo_pen),
           enframe(non_comp_trac_second_wo_pen_normal),
           enframe(non_comp_trac_second_wo_pen_wo_pen),
           enframe(non_comp_slc_second),
          enframe(non_comp_slc_second_wo_pen),
           enframe(non_comp_slc_second_wo_pen_normal),
           enframe(non_comp_slc_second_wo_pen_wo_pen)) %>%
  mutate(val = 1) 

p_non_comp <- df_joint_selected %>%
  expand(name, value) %>%
  left_join(df_joint_selected, by = c("name", "value")) %>%
  mutate(val = replace_na(val, 0)) %>%
  mutate(val = factor(val, levels = c(1, 0),
                      labels = c("Yes", "No"))) %>%
  mutate(value = str_replace_all(value, "allergic_to_i_have_no_food_allergies_that_i_know_of", "no_food_allergies")) %>%
  mutate(name = str_replace_all(name, "_", " ")) %>%
  mutate(value = str_replace_all(value, "_", " ")) %>%
  mutate(value = str_replace_all(value, "sex", "female")) %>%
  ggplot(aes(name, value, fill= val)) + 
  geom_tile(colour = "gray50") +
  scale_fill_manual(values = c("black", "white")) +
  coord_equal(expand = 0) +
  theme_bw() +
  theme(panel.grid.major = element_blank(),
        axis.text.x = element_text(angle = 45, hjust = 1)) +
  xlab("") +
  ylab("") + 
  guides(fill = guide_legend(title = "Selected?"))

ggsave(filename = "../AGP_age/results/non_comp_selected.png", plot = p_non_comp)
p_non_comp
```

```{r}
tax_level <- c("Phylum", "Class", "Order", "Family", "Genus",
  "Species", "OTU")
tax_levels <- factor(tax_level, level = tax_level)

comp_trac <- names(nz[!get_non_compositional_trac(nz)]) %>%
  str_replace(., "'", "") %>%
  str_count(., "::") %>%
  {table(tax_levels[.])}

comp_trac_wo_pen <- names(nz_wo_pen[!get_non_compositional_trac(nz_wo_pen)]) %>%
  str_replace(., "'", "") %>%
  str_count(., "::") %>%
  {table(tax_levels[.])}

comp_trac_baseline <- names(nz_baseline[!get_non_compositional_trac(nz_baseline)]) %>%
  str_replace(., "'", "") %>%
  str_count(., "::") %>%
  {table(tax_levels[.])}

comp_slc <- names(nz_slc[!get_non_compositional_slc(nz_slc)]) %>%
  str_count(., "::") %>%
  {table(tax_levels[. + 7])}
comp_slc_wo_pen <- names(nz_slc_wo_pen[!get_non_compositional_slc(nz_slc_wo_pen)])  %>%
  str_count(., "::") %>%
  {table(tax_levels[. + 7])}
comp_slc_baseline <- names(nz_slc_baseline[!get_non_compositional_slc(nz_slc_baseline)]) %>%
  str_count(., "::") %>%
  {table(tax_levels[. + 7])}


tax_df <- bind_rows(comp_trac,
                    comp_trac_wo_pen,
                    comp_trac_baseline,
                    comp_slc,
                    comp_slc_wo_pen,
                    comp_slc_baseline) %>%
  as_tibble() %>%
  as.data.frame()

tax_df <- tibble(Method = c("trac", "trac (w = 0.5)", "trac only microbiome",
                            "slc", "slc (w = 0.5)", "slc only microbiome")) %>%
  bind_cols(., tax_df) %>%
  bind_cols(., tibble(Meta = c(length(non_comp_trac),
                               length(non_comp_trac_wo_pen),
                               0,
                               length(non_comp_slc),
                               length(non_comp_slc_wp_pen),
                               0)))


tax_df %>%
  kbl("latex") %>%
  save_kable(
    "../AGP_age/results/summary_selected_levels.tex") 
tax_df
```

## Tables selected log-ratios

### trac

```{r}
enframe(nz_second) %>%
  arrange(desc(abs(value))) %>%
  slice_head(n = 10) %>%
  kbl("latex") %>%
  save_kable(
    "../AGP_age/results/trac_two_normal_normal.tex") 

# Second Stage trac normal normal
kbl(enframe(nz_second) %>%
  arrange(desc(abs(value))))
```
```{r}
# Second Stage trac normal less penalization
enframe(nz_second_wo_pen) %>%
  arrange(desc(abs(value))) %>%
  slice_head(n = 10) %>%
  kbl("latex") %>%
  save_kable(
    "../AGP_age/results/trac_two_normal_wo_pen.tex") 

kbl(enframe(nz_second_wo_pen) %>%
  arrange(desc(abs(value))))
```


```{r}
# Second Stage trac less penalization normal
enframe(nz_second_wo_pen_normal) %>%
  arrange(desc(abs(value))) %>%
  slice_head(n = 10) %>%
  kbl("latex") %>%
  save_kable(
    "../AGP_age/results/trac_two_wo_pen_normal.tex") 

kbl(enframe(nz_second_wo_pen_normal) %>%
  arrange(desc(abs(value))))
```

```{r}
# Second Stage trac less less penalization
enframe(nz_second_wo_pen_wo_pen) %>%
  arrange(desc(abs(value))) %>%
  slice_head(n = 10) %>%
  kbl("latex") %>%
  save_kable(
    "../AGP_age/results/trac_two_wo_pen_wo_pen.tex") 

kbl(enframe(nz_second_wo_pen_wo_pen) %>%
  arrange(desc(abs(value))))
```

### SLC

```{r}
enframe(nz_second_slc) %>%
  arrange(desc(abs(value))) %>%
  slice_head(n = 10) %>%
  kbl("latex") %>%
  save_kable(
    "../AGP_age/results/slc_two_normal_normal.tex") 

# Second Stage trac normal normal
kbl(enframe(nz_second_slc) %>%
  arrange(desc(abs(value))))
```

```{r}
# Second Stage trac normal less penalization
enframe(nz_second_slc_wo_pen) %>%
  arrange(desc(abs(value))) %>%
  slice_head(n = 10) %>%
  kbl("latex") %>%
  save_kable(
    "../AGP_age/results/slc_two_normal_wo_pen.tex") 

kbl(enframe(nz_second_slc_wo_pen) %>%
  arrange(desc(abs(value))))
```


```{r}
# Second Stage trac less penalization normal
enframe(nz_second_slc_wo_pen_normal) %>%
  arrange(desc(abs(value))) %>%
  slice_head(n = 10) %>%
  kbl("latex") %>%
  save_kable(
    "../AGP_age/results/slc_two_wo_pen_normal.tex") 

kbl(enframe(nz_second_slc_wo_pen_normal) %>%
  arrange(desc(abs(value))))
```

```{r}
# Second Stage trac less less penalization
enframe(nz_second_slc_wo_pen_wo_pen) %>%
  arrange(desc(abs(value))) %>%
  slice_head(n = 10) %>%
  kbl("latex") %>%
  save_kable(
    "../AGP_age/results/slc_two_wo_pen_wo_pen.tex") 

kbl(enframe(nz_second_slc_wo_pen_wo_pen) %>%
  arrange(desc(abs(value))))
```

### Baseline

```{r}
# Second Stage trac baseline
enframe(nz_second_baseline) %>%
  arrange(desc(abs(value))) %>%
  slice_head(n = 10) %>%
  kbl("latex") %>%
  save_kable(
    "../AGP_age/results/trac_two_baseline.tex") 

kbl(enframe(nz_second_baseline) %>%
  arrange(desc(abs(value))))
```


```{r}
# Second Stage slc baseline
enframe(nz_second_slc_baseline) %>%
  arrange(desc(abs(value))) %>%
  slice_head(n = 10) %>%
  kbl("latex") %>%
  save_kable(
    "../AGP_age/results/slc_two_baseline.tex") 

kbl(enframe(nz_second_slc_baseline) %>%
  arrange(desc(abs(value))))
```

